---
title: "Predictions for the 2023 ICC Cricket World Cup"
output: pdf_document
date: "2023-11-30"
---

```{r}
library(ggplot2)
library(caTools)
library(randomForest)
library(dplyr)
library(fastDummies)
library(tidyverse)
library(caret)
library(pROC)
library(gplots)
library(corrplot)
library(ggcorrplot)
```


```{r}
#Reading the dataset
World_cup <- read.csv("World_cup_2023.csv")
results <- read.csv("results.csv")

World_cup
```

No.of titles won by each teams

```{r}
ggplot(World_cup, aes(x = Team_name, y = Titles, fill = Team_name)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = rainbow(nrow(World_cup))) +  # Use rainbow colors
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
  labs(title = "Team Titles in World Cup",
       x = "Team Name",
       y = "Titles")

```


Win percentage in ODI by each team

```{r}
ggplot(World_cup, aes(x = Team_name, y = Win_percentage_ODI, fill = Team_name)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = rainbow(nrow(World_cup))) +  # Use rainbow colors
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
  labs(title = "Team Titles in World Cup",
       x = "Team Name",
       y = "Titles")

```


No.of matches won in world cup by each team


```{r}
ggplot(World_cup, aes(x = Team_name, y = WC_match_won, fill = Team_name)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = rainbow(nrow(World_cup))) +  # Use rainbow colors
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
  labs(title = "Team Titles in World Cup",
       x = "Team Name",
       y = "Titles")
```


Recent ICC ODI rating

```{r}

ggplot(World_cup, aes(x = Team_name, y = Rating, fill = Team_name)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = rainbow(nrow(World_cup))) +  # Use rainbow colors
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
  labs(title = "Team Titles in World Cup",
       x = "Team Name",
       y = "Titles")

```



```{r}
head(results)
```

```{r}
#Removing the Match abandoned and No result data from Winner column
results <- results[results$Winner != 'Match abandoned', ]
results <- results[results$Winner != 'No result', ]

```



Stats of top five teams in the ODI world cup 2023

Stats of Indian team


```{r}

#Filtering India
df <- results[results$Team_1 == 'India' | results$Team_2 == 'India', ]
India <- df

head(India)


```


```{r}

India_win <- India[India$Winner == 'India', ]
head(India_win)

```


No.of wins in ODIs against other teams


```{r}
excluded_value <- 'India'
filtered_df <- India_win[India_win$Team_2 != excluded_value, ]
value_counts <- table(filtered_df$Team_2)
print(value_counts)

```


```{r}

excluded_value <- 'India'

filtered_df <- India_win[India_win$Team_1 != excluded_value, ]

value_counts <- table(filtered_df$Team_1)

print(value_counts)

```


Displaying Team India's wins against other teams using bar graph

```{r}

exclude <- 'India'

filtered_data <- India_win[India_win$Team_2 != exclude, ]

top_opponents <- head(names(sort(table(filtered_data$Team_2), decreasing = TRUE)), 5)
top_opponents_counts <- as.numeric(table(filtered_data$Team_2)[top_opponents])

barplot(top_opponents_counts, names.arg = top_opponents, col = "blue", main = "Top 5 Opponents Faced by India", xlab = "Opponent", ylab = "Number of Matches")

```


```{r}
exclude <- 'India'

filtered_data <- India_win[India_win$Team_1 != exclude, ]

top_opponents <- head(names(sort(table(filtered_data$Team_1), decreasing = TRUE)), 5)
top_opponents_counts <- as.numeric(table(filtered_data$Team_1)[top_opponents])

barplot(top_opponents_counts, names.arg = top_opponents, col = "blue", main = "Top 5 Opponents Faced by India", xlab = "Opponent", ylab = "Number of Matches")

```


Win Percentage of India Against Each Team


```{r}
# Number of wins against each team
team_win_counts <- c(
  Australia = 54,
  `New Zealand` = 58,
  `South Africa` = 37,
  Pakistan = 55,
  `Sri Lanka` = 93,
  Bangladesh = 30,
  England = 57,
  Netherlands = 2,
  Afghanistan = 3
)

total_matches <- sum(team_win_counts)

win_percentages <- (team_win_counts / total_matches) * 100

# Pie chart
pie(win_percentages, labels = names(win_percentages), col = rainbow(length(win_percentages)), 
    main = "Win Percentage of India Against Each Team", cex.main = 0.8, clockwise = TRUE)

legend("topright", legend = paste(names(win_percentages), sprintf("(%.1f%%)", win_percentages), sep = " "), cex = 0.8, fill = rainbow(length(win_percentages)))

```



Win Percentage of India in the ODI world cup

```{r}
# Number of wins against each team in the ODI world cup
team_win_counts_wc_ind <- c(
  Australia = 4,
  `New Zealand` = 3,
  `South Africa` = 2,
  Pakistan = 7,
  `Sri Lanka` = 5,
  Bangladesh = 3,
  England = 3,
  Netherlands = 2,
  Afghanistan = 2
)

total_matches_wc_ind <- sum(team_win_counts_wc_ind)

win_percentages_wc_ind <- (team_win_counts_wc_ind / total_matches_wc_ind) * 100

pie(win_percentages_wc_ind, labels = names(win_percentages_wc_ind), col = rainbow(length(win_percentages_wc_ind)), 
    main = "Win Percentage of India in the ODI World Cup", cex.main = 0.8, clockwise = TRUE)

legend("topright", legend = paste(names(win_percentages_wc_ind), sprintf("(%.1f%%)", win_percentages_wc_ind), sep = " "), cex = 0.8, fill = rainbow(length(win_percentages_wc_ind)))

```


Stats of Australian team

```{r}
# Filter the Australia Data
df1 <- results[results$Team_1 == 'Australia' | results$Team_2 == 'Australia', ]

Australia <- df1

head(Australia)

```


```{r}
Australia_win <- Australia[Australia$Winner == 'Australia', ]
head(Australia_win)

```

No.of wins for Australia against other teams

```{r}
excluded_value <- 'Australia'
filtered_df <- Australia_win[Australia_win$Team_2 != excluded_value, ]

value_counts <- table(filtered_df$Team_2)

print(value_counts)

```


```{r}
excluded_value <- 'Australia'

filtered_df <- Australia_win[Australia_win$Team_1 != excluded_value, ]

value_counts <- table(filtered_df$Team_1)

print(value_counts)
```

Displaying Team Australia's wins against other teams using bar graph

```{r}
exclude <- 'Australia'

filtered_data <- Australia_win[Australia_win$Team_2 != exclude, ]

top_opponents <- head(names(sort(table(filtered_data$Team_2), decreasing = TRUE)), 5)
top_opponents_counts <- as.numeric(table(filtered_data$Team_2)[top_opponents])

barplot(top_opponents_counts, names.arg = top_opponents, col = "darkgreen", main = "Top 5 Opponents Faced by Australia", xlab = "Opponent", ylab = "Number of Matches")
```


```{r}
exclude <- 'Australia'

filtered_data <- Australia_win[Australia_win$Team_1 != exclude, ]

top_opponents <- head(names(sort(table(filtered_data$Team_1), decreasing = TRUE)), 5)
top_opponents_counts <- as.numeric(table(filtered_data$Team_1)[top_opponents])

barplot(top_opponents_counts, names.arg = top_opponents, col = "darkgreen", main = "Top 5 Opponents Faced by Australia", xlab = "Opponent", ylab = "Number of Matches")
```


Win Percentage of Australia Against Each Team

```{r}
# Number of wins against each team
team_win_counts <- c(
  'India'= 82,
  'New Zealand'= 95,
  'South Africa '= 48,
  'Pakistan'= 69,
  'Sri Lanka'= 63,
  'Bangladesh'= 19,
  'England'= 87,
  'Netherlands'= 2,
  'Afghanistan'= 3
)

total_matches <- sum(team_win_counts)

win_percentages <- (team_win_counts / total_matches) * 100

pie(win_percentages, labels = names(win_percentages), col = rainbow(length(win_percentages)), 
    main = "Win Percentage of Australia Against Each Team", cex.main = 0.8, clockwise = TRUE)


legend("topright", legend = paste(names(win_percentages), sprintf("(%.1f%%)", win_percentages), sep = " "), cex = 0.8, fill = rainbow(length(win_percentages)))
```


Win Percentage of Australia in the ODI world cup

```{r}
# Number of wins against each team in the ODI world cup
team_win_counts_wc_ind <- c(
  'India'= 8,
  'New Zealand'= 8,
  'South Africa '= 3,
  'Pakistan'= 6,
  'Sri Lanka'= 8,
  'Bangladesh'= 3,
  'England'= 6,
  'Netherlands'= 2,
  'Afghanistan'= 2
)

total_matches_wc_ind <- sum(team_win_counts_wc_ind)

win_percentages_wc_ind <- (team_win_counts_wc_ind / total_matches_wc_ind) * 100

pie(win_percentages_wc_ind, labels = names(win_percentages_wc_ind), col = rainbow(length(win_percentages_wc_ind)), 
    main = "Win Percentage of Australia in the ODI World Cup", cex.main = 0.8, clockwise = TRUE)

legend("topright", legend = paste(names(win_percentages_wc_ind), sprintf("(%.1f%%)", win_percentages_wc_ind), sep = " "), cex = 0.8, fill = rainbow(length(win_percentages_wc_ind)))
```



Stats of Pakistan team

```{r}
#Filter the Pakistan Data
df2 <- results[results$Team_1 == 'Pakistan' | results$Team_2 == 'Pakistan', ]
Pakistan <- df2

head(Pakistan)
```


```{r}
Pakistan_win <- Pakistan[Pakistan$Winner == 'Pakistan', ]
head(Pakistan_win)

```


No.of wins for Pakistan against other teams

```{r}
excluded_value <- 'Pakistan'

filtered_df <- Pakistan_win[Pakistan_win$Team_2 != excluded_value, ]

value_counts <- table(filtered_df$Team_2)

print(value_counts)
```


```{r}
excluded_value <- 'Pakistan'

filtered_df <- Pakistan_win[Pakistan_win$Team_1 != excluded_value, ]

value_counts <- table(filtered_df$Team_1)

print(value_counts)

```


Displaying Team Pakistan's wins against other teams using bar graph

```{r}
exclude <- 'Pakistan'

filtered_data <- Pakistan_win[Pakistan_win$Team_2 != exclude, ]

top_opponents <- head(names(sort(table(filtered_data$Team_2), decreasing = TRUE)), 5)
top_opponents_counts <- as.numeric(table(filtered_data$Team_2)[top_opponents])

barplot(top_opponents_counts, names.arg = top_opponents, col = "darkblue", main = "Top 5 Opponents Faced by Pakistan", xlab = "Opponent", ylab = "Number of Matches")
```


```{r}
exclude <- 'Pakistan'

filtered_data <- Pakistan_win[Pakistan_win$Team_1 != exclude, ]

top_opponents <- head(names(sort(table(filtered_data$Team_1), decreasing = TRUE)), 5)
top_opponents_counts <- as.numeric(table(filtered_data$Team_1)[top_opponents])

barplot(top_opponents_counts, names.arg = top_opponents, col = "darkblue", main = "Top 5 Opponents Faced by Pakistan", xlab = "Opponent", ylab = "Number of Matches")
```


Win Percentage of Pakistan Against Each Team

```{r}
# Number of wins against each team
team_win_counts <- c(
  'India'= 73,
  'New Zealand'= 60,
  'South Africa '= 30,
  'Australia'= 34,
  'Sri Lanka'= 92,
  'Bangladesh'= 32,
  'England'= 32,
  'Netherlands'= 3,
  'Afghanistan'= 7
)

total_matches <- sum(team_win_counts)

win_percentages <- (team_win_counts / total_matches) * 100

pie(win_percentages, labels = names(win_percentages), col = rainbow(length(win_percentages)), 
    main = "Win Percentage of Pakistan Against Each Team", cex.main = 0.8, clockwise = TRUE)


legend("topright", legend = paste(names(win_percentages), sprintf("(%.1f%%)", win_percentages), sep = " "), cex = 0.8, fill = rainbow(length(win_percentages)))
```


Win Percentage of Pakistan in the ODI world cup

```{r}
# Number of wins against each team in the ODI world cup
team_win_counts_wc_ind <- c(
  'India'= 0,
  'New Zealand'= 7,
  'South Africa '= 2,
  'Australia'= 4,
  'Sri Lanka'= 7,
  'Bangladesh'= 1,
  'England'= 5,
  'Afghanistan'= 1
)

total_matches_wc_ind <- sum(team_win_counts_wc_ind)

win_percentages_wc_ind <- (team_win_counts_wc_ind / total_matches_wc_ind) * 100

pie(win_percentages_wc_ind, labels = names(win_percentages_wc_ind), col = rainbow(length(win_percentages_wc_ind)), 
    main = "Win Percentage of Pakistan in the ODI World Cup", cex.main = 0.8, clockwise = TRUE)

legend("topright", legend = paste(names(win_percentages_wc_ind), sprintf("(%.1f%%)", win_percentages_wc_ind), sep = " "), cex = 0.8, fill = rainbow(length(win_percentages_wc_ind)))
```


Stats of New Zealand team

```{r}
# Filter the New Zealand data
df3 <- results[results$Team_1 == 'New Zealand' | results$Team_2 == 'New Zealand', ]
New_Zealand <- df3

head(New_Zealand)
```


```{r}
New_Zealand_win <- New_Zealand[New_Zealand$Winner == 'New Zealand', ]
head(New_Zealand_win)

```

No.of wins for New Zealand in ODIs against other teams

```{r}
excluded_value <- 'New Zealand'

filtered_df <- New_Zealand_win[New_Zealand_win$Team_2 != excluded_value, ]

value_counts <- table(filtered_df$Team_2)

print(value_counts)
```

```{r}
excluded_value <- 'New Zealand'

filtered_df <- New_Zealand_win[New_Zealand_win$Team_1 != excluded_value, ]

value_counts <- table(filtered_df$Team_1)

print(value_counts)
```

Displaying Team New Zealand's wins against other teams using bar graph

```{r}
exclude <- 'New Zealand'

filtered_data <- New_Zealand_win[New_Zealand_win$Team_2 != exclude, ]

top_opponents <- head(names(sort(table(filtered_data$Team_2), decreasing = TRUE)), 5)
top_opponents_counts <- as.numeric(table(filtered_data$Team_2)[top_opponents])

barplot(top_opponents_counts, names.arg = top_opponents, col = "yellow", main = "Top 5 Opponents Faced by New Zealand", xlab = "Opponent", ylab = "Number of Matches")
```


```{r}
exclude <- 'New Zealand'

filtered_data <- New_Zealand_win[New_Zealand_win$Team_1 != exclude, ]

top_opponents <- head(names(sort(table(filtered_data$Team_1), decreasing = TRUE)), 5)
top_opponents_counts <- as.numeric(table(filtered_data$Team_1)[top_opponents])

barplot(top_opponents_counts, names.arg = top_opponents, col = "yellow", main = "Top 5 Opponents Faced by New Zealand", xlab = "Opponent", ylab = "Number of Matches")
```



Win Percentage of New Zealand Against Each Team


```{r}
# Number of wins against each team
team_win_counts <- c(
  'Australia'= 39,
  'India'= 50,
  'South Africa '= 26,
  'Pakistan'= 50,
  'Sri Lanka'= 49,
  'Bangladesh'= 28,
  'England'= 42,
  'Netherlands'= 4,
  'Afghanistan'= 2
)

total_matches <- sum(team_win_counts)

win_percentages <- (team_win_counts / total_matches) * 100

pie(win_percentages, labels = names(win_percentages), col = rainbow(length(win_percentages)), 
    main = "Win Percentage of New Zealand Against Each Team", cex.main = 0.8, clockwise = TRUE)


legend("topright", legend = paste(names(win_percentages), sprintf("(%.1f%%)", win_percentages), sep = " "), cex = 0.8, fill = rainbow(length(win_percentages)))
```


Win Percentage of New Zealand in the ODI world cup

```{r}
# Number of wins against each team in the ODI world cup
team_win_counts_wc_ind <- c(
  'Australia'= 3,
  'India'= 5,
  'South Africa '= 5,
  'Pakistan'= 2,
  'Sri Lanka'= 5,
  'Bangladesh'= 5,
  'England'= 5,
  'Netherlands'= 4,
  'Afghanistan'= 2
)

total_matches_wc_ind <- sum(team_win_counts_wc_ind)

win_percentages_wc_ind <- (team_win_counts_wc_ind / total_matches_wc_ind) * 100

pie(win_percentages_wc_ind, labels = names(win_percentages_wc_ind), col = rainbow(length(win_percentages_wc_ind)), 
    main = "Win Percentage of New Zealand in the ODI World Cup", cex.main = 0.8, clockwise = TRUE)

legend("topright", legend = paste(names(win_percentages_wc_ind), sprintf("(%.1f%%)", win_percentages_wc_ind), sep = " "), cex = 0.8, fill = rainbow(length(win_percentages_wc_ind)))
```


Stats of England team

```{r}
# Filter the England data
df4 <- results[results$Team_1 == 'England' | results$Team_2 == 'England', ]

England <- df4

head(England)
```

```{r}
England_win <- England[England$Winner == 'England', ]
head(England_win)

```

No.of wins for England against other teams

```{r}
excluded_value <- 'England'

filtered_df <- England_win[England_win$Team_2 != excluded_value, ]

value_counts <- table(filtered_df$Team_2)

print(value_counts)
```


```{r}
excluded_value <- 'England'

filtered_df <- England_win[England_win$Team_1 != excluded_value, ]

value_counts <- table(filtered_df$Team_1)

print(value_counts)
```


Displaying Team England's wins against other teams using bar graph

```{r}
exclude <- 'England'

filtered_data <- England_win[England_win$Team_2 != exclude, ]

top_opponents <- head(names(sort(table(filtered_data$Team_2), decreasing = TRUE)), 5)
top_opponents_counts <- as.numeric(table(filtered_data$Team_2)[top_opponents])

barplot(top_opponents_counts, names.arg = top_opponents, col = "purple", main = "Top 5 Opponents Faced by England", xlab = "Opponent", ylab = "Number of Matches")
```


```{r}
exclude <- 'England'

filtered_data <- England_win[England_win$Team_1 != exclude, ]

top_opponents <- head(names(sort(table(filtered_data$Team_1), decreasing = TRUE)), 5)
top_opponents_counts <- as.numeric(table(filtered_data$Team_1)[top_opponents])

barplot(top_opponents_counts, names.arg = top_opponents, col = "purple", main = "Top 5 Opponents Faced by England", xlab = "Opponent", ylab = "Number of Matches")
```


Win Percentage of England Against Each Team

```{r}
# Number of wins against each team
team_win_counts <- c(
  'India'= 44,
  'New Zealand'= 42,
  'South Africa '= 29,
  'Australia'= 63,
  'Sri Lanka'= 38,
  'Bangladesh'= 19,
  'Pakistan'= 56,
  'Netherlands'= 6,
  'Afghanistan'= 2
)

total_matches <- sum(team_win_counts)

win_percentages <- (team_win_counts / total_matches) * 100

pie(win_percentages, labels = names(win_percentages), col = rainbow(length(win_percentages)), 
    main = "Win Percentage of England Against Each Team", cex.main = 0.8, clockwise = TRUE)


legend("topright", legend = paste(names(win_percentages), sprintf("(%.1f%%)", win_percentages), sep = " "), cex = 0.8, fill = rainbow(length(win_percentages)))
```


```{r}
# Number of wins against each team in the ODI world cup
team_win_counts_wc_ind <- c(
  'Australia'= 3,
  'India'= 4,
  'South Africa' = 4,
  'Pakistan'= 4,
  'Sri Lanka'= 6,
  'Bangladesh'= 5,
  'New Zealand'= 5,
  'Netherlands'= 4,
  'Afghanistan'= 2
)

total_matches_wc_ind <- sum(team_win_counts_wc_ind)

win_percentages_wc_ind <- (team_win_counts_wc_ind / total_matches_wc_ind) * 100

pie(win_percentages_wc_ind, labels = names(win_percentages_wc_ind), col = rainbow(length(win_percentages_wc_ind)), 
    main = "Win Percentage of England in the ODI World Cup", cex.main = 0.8, clockwise = TRUE)

legend("topright", legend = paste(names(win_percentages_wc_ind), sprintf("(%.1f%%)", win_percentages_wc_ind), sep = " "), cex = 0.8, fill = rainbow(length(win_percentages_wc_ind)))
```

```{r}
modified_WorldCup_data = World_cup
modified_WorldCup_data$World_cup_winner <- ifelse(modified_WorldCup_data$World_cup_winner == "Yes", 1, 0)
modified_WorldCup_data = subset(modified_WorldCup_data, select = -c(Team_name))

correlation_matrix <- cor(modified_WorldCup_data)

#corrplot(correlation_matrix, method = "color", type = "upper", addCoef.col = "white")
ggcorrplot(correlation_matrix,
           hc.order = TRUE,
           type = "lower",
           method = "circle",
           outline.color = "white",
           col = c("yellow", "orange", "green"),
           pch = c(19, 15, 17),
           tl.cex = 1.2,
           tl.col = "black") +
  labs(title = "Correlation Matrix of ICC World Cup Dataset") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
        axis.text.x = element_text(angle = 90, hjust = 1))
```




```{r}
# cricket teams participating in the world cup
worldcup_teams <- c('England', ' South Africa', '', 'West Indies',
                    'Pakistan', 'New Zealand', 'Sri Lanka', 'Afghanistan',
                    'Australia', 'Bangladesh', 'India')

# Filtering data involving only teams in the 'worldcup_teams' list
df_teams_1 <- results[results$Team_1 %in% worldcup_teams, ]
df_teams_2 <- results[results$Team_2 %in% worldcup_teams, ]

df_teams <- rbind(df_teams_1, df_teams_2)
nrow(df_teams)

```


```{r}
head(df_teams)
```


```{r}
df_teams_2019 <- subset(df_teams, select = -c(Date, Margin, Ground))

head(df_teams_2019)

```



Feature selection

```{r}

df_teams_2019 <- df_teams_2019[ , !(names(df_teams_2019) %in% c("winning_team"))]
rownames(df_teams_2019) <- NULL

df_teams_2019$winning_team <- ifelse(df_teams_2019$Winner == df_teams_2019$Team_1, 1, NA)

df_teams_2019$winning_team[is.na(df_teams_2019$winning_team) & df_teams_2019$Winner == df_teams_2019$Team_2] <- 2

df_teams_2019 <- subset(df_teams_2019, select = -c(winning_team))

head(df_teams_2019)


```


Training and testing the data

```{r}
final <- data.frame(model.matrix(~ Team_1 + Team_2 - 1, data = df_teams_2019), df_teams_2019)

X <- final[, !(names(final) %in% c("Winner"))]

y <- final$Winner

set.seed(42)

# training and testing partition
split <- sample.split(y, SplitRatio = 0.8)

# splitting training and testing sets
X_train <- X[split, ]
X_test <- X[!split, ]
y_train <- y[split]
y_test <- y[!split]


```



```{r}
head(final)
```




Applying Random forest classifier algorithm



```{r}
y_train <- as.factor(y_train)
y_test <- as.factor(y_test)

rf <- randomForest(y_train ~ ., data = cbind(X_train, y_train), ntree = 100, mtry = 20, nodesize = 1, importance = TRUE, seed = 0)

pred <- predict(rf, newdata = X_test)

levels_pred <- levels(pred)
levels_y_test <- levels(y_test)
common_levels <- union(levels_pred, levels_y_test)

pred <- factor(pred, levels = common_levels)
y_test <- factor(y_test, levels = common_levels)

score <- sum(predict(rf) == y_train) / length(y_train)

score2 <- sum(pred == y_test) / length(y_test)

cat("Training set accuracy:", sprintf("%.3f\n", score))

cat("Test set accuracy:", sprintf("%.3f\n", score2))


```



Evaluating the best model on test data


```{r}
ranking <- read.csv('Icc_ranking.csv')

fixtures <- read.csv('Fixtures.csv')

pred_set <- list()
```


```{r}

fixtures <- fixtures %>%
  mutate(first_position = ranking$Team_ranking[match(Team_1, ranking$Team_name)])

fixtures <- fixtures %>%
  mutate(second_position = ranking$Team_ranking[match(Team_2, ranking$Team_name)])
fixtures <- head(fixtures, 45)
head(fixtures)


```


```{r}
fixtures$first_position <- ifelse(is.na(fixtures$first_position),
                                  mean(fixtures$first_position, na.rm = TRUE),
                                  fixtures$first_position)

fixtures$second_position <- ifelse(is.na(fixtures$second_position),
                                   mean(fixtures$second_position, na.rm = TRUE),
                                   fixtures$second_position)

head(fixtures)

```


```{r}
# Predicting winning teams 
pred_set <- list()

for (i in 1:nrow(fixtures)) {
  if (fixtures$first_position[i] < fixtures$second_position[i]) {
    pred_set[[i]] <- data.frame(Team_1 = fixtures$Team_1[i],
                                 Team_2 = fixtures$Team_2[i],
                                 winning_team = NA)
  } else {
    pred_set[[i]] <- data.frame(Team_1 = fixtures$Team_2[i],
                                 Team_2 = fixtures$Team_1[i],
                                 winning_team = NA)
  }
}

pred_set <- do.call(rbind, pred_set)

backup_pred_set <- pred_set

head(pred_set)

```


```{r}
pred_set <- dummy_cols(pred_set, select_columns = c("Team_1", "Team_2"), remove_selected_columns = TRUE)

pred_set <- as.data.frame(pred_set)

missing_cols <- setdiff(names(final), names(pred_set))

for (c in missing_cols) {
  pred_set[[c]] <- 0
}

pred_set <- pred_set[, names(final)]

pred_set <- pred_set[, !names(pred_set) %in% 'Winner']

head(pred_set)

```


Interpret the model results

```{r}
# Predictions
predictions <- predict(rf, newdata = pred_set, type = "response")

for (i in 1:nrow(fixtures)) {
  cat(backup_pred_set[i, "Team_2"], " vs ", backup_pred_set[i, "Team_1"], "\n")
  
  if (predictions[i] == 1) {
    cat("Winner: ", backup_pred_set[i, "Team_2"], "\n")
  } else {
    cat("Winner: ", backup_pred_set[i, "Team_1"], "\n")
  }
  
  cat("\n")
}

```


```{r}

semi <- list(list('Pakistan', 'Australia'), list('England', 'India'))

print(semi)

```


```{r}
#function to predict Winner
clean_and_predict <- function(matches, ranking, final, rf) {
  positions <- c()

  for (match in matches) {
    positions <- c(positions, ranking$Team_ranking[ranking$Team_name == match[1]])
    positions <- c(positions, ranking$Team_ranking[ranking$Team_name == match[2]])
  }
  
  pred_set <- list()
  
  i <- 1
  j <- 1
  
  while (i <= length(positions)) {
    dict1 <- list(
      Team_1 = c(),
      Team_2 = c()
    )
    
    if (positions[i] < positions[i + 1]) {
      dict1$Team_1 <- c(dict1$Team_1, matches[[j]][[1]])
      dict1$Team_2 <- c(dict1$Team_2, matches[[j]][[2]])
    } else {
      dict1$Team_1 <- c(dict1$Team_1, matches[[j]][[2]])
      dict1$Team_2 <- c(dict1$Team_2, matches[[j]][[1]])
    }
    pred_set <- c(pred_set, list(dict1))
    i <- i + 2
    j <- j + 1
  }
  
  pred_set_df <- do.call(rbind, lapply(pred_set, data.frame, stringsAsFactors = FALSE))
  
  backup_pred_set <- pred_set_df

  pred_set_df <- dummy_cols(pred_set_df, select_columns = c("Team_1", "Team_2"), remove_selected_columns = TRUE)
  
  missing_cols2 <- setdiff(names(final), names(pred_set_df))
  
  for (c in missing_cols2) {
    pred_set_df[[c]] <- 0
  }
  
  pred_set_df <- pred_set_df %>%
    dplyr::select(names(final))

  pred_set_df <- pred_set_df[, !(names(pred_set_df) %in% c("Winner"))]
  
  predictions <- predict(rf, newdata = pred_set_df)

  for (i in 1:nrow(pred_set_df)) {
    cat(paste(backup_pred_set[i, "Team_2"], " vs ", backup_pred_set[i, "Team_1"], "\n"))
    
    if (predictions[i] == 1) {
      cat("Winner: ", backup_pred_set[i, "Team_2"], "\n")
    } else {
      cat("Winner: ", backup_pred_set[i, "Team_1"], "\n")
    }
    
    cat("\n")
  }
  
}

```


```{r}
clean_and_predict(semi, ranking, final, rf)
```

```{r}
finals = list(list('Australia', 'India'))
```


```{r}
clean_and_predict(finals, ranking, final, rf)
```


```{r}
# Confusion Matrix Creation
Y_predicted <- factor(pred)
y_test <- factor(y_test)

common_levels <- union(levels(Y_predicted), levels(y_test))

Y_predicted <- factor(Y_predicted, levels = common_levels)
y_test <- factor(y_test, levels = common_levels)

conf_matrix <- confusionMatrix(Y_predicted, y_test)


```


```{r}

heatmap(conf_matrix$table, 
        Colv = NA, Rowv = NA, 
        margins=c(10,10),
        col = colorRampPalette(c("white", "steelblue"))(100),
        main = "Confusion Matrix for Random Forest",
        xlab = "Predicted Values",
        ylab = "Test Values")

```


